\SW: Crowd-sourcing the Discovery of Gravitational Lenses

### 2014-05-14 doc/sw-system.tex
Description of CI, screenshot fig

Systematic detection of rare astronomical objects by such ``crowd-sourced''
visual inspection has recently been achieved by the online citizen science
project PlanetHunters \citep{SchwambEtal2012}. PlanetHunters was designed to
enable the discovery of transiting exoplanets in data taken by the Kepler
satellite; a community of N inspectors found, after each undergoing a small
amount of training, N new exoplanet candidates by visual inspection of the
Kepler lightcurves that were presented in a custom web-based classification
interface. The older Galaxy Zoo morphological classification project
\citep{LintottEtal2008} has also enabled the discovery of rare objects, via its
flexible inspection interface and discussion forum \citep{LintottEtal2009}. Indeed, several of us
(AV,EB,CC,TJ,CM,LW) were active in an informal Galaxy Zoo gravitational lens
search, an experience which led to the present hypothesis that a systematic
online visual lens search could be successful.

In this paper, we describe the \sw website, an online system that enables 
crowd-sourced gravitational lens detection.  In a companion paper we will
present the new gravitational lenses discovered in our first experimental lens
search, and begin to investigate the differences between lens detections made
in \sw and those made with automated techniques. Here though, we try to answer
the following questions:

@@

This paper is organised as follows. 
In \Sref{sec:design} we introduce the \sw classification interface,  explain
how we use training images, and describe our two stage candidate selection
strategy. We then briefly introduce, in \Sref{sec:data} the particular dataset
used in the first experimental tests of the \sw system, and how we prepared
the images prior to displaying them in the web interface. In \Sref{sec:swap}
we describe our methodology for interpreting the classifications made by the
volunteers, and then present the results of system performance tests 
made on the training images in \Sref{sec:results}.
% In this section, we also investigate the properties of the crowd, to
% understand where the information is coming from.
We discuss the implications of our results for future lens searches in
\Sref{sec:discuss} and draw conclusions in \Sref{sec:conclude}.

@@

Unfamiliar objects: need to learn what lenses look like, fast. Rare
objects: need to be able to reject rapidly, and get through sample.
Confusion with non-lenses: further filtering after advanced training, and 
scientific discussion in Talk.

Intro to Classification Interface. Basic description of site.

@@

>>>>

Systematic detection of rare astronomical objects by such ``crowd-sourced''
visual inspection has recently been achieved by the online citizen science
project PlanetHunters \citep{SchwambEtal2012}. PlanetHunters was designed to
enable the discovery of transiting exoplanets in data taken by the Kepler
satellite; a community of N inspectors from the general public found, after
each undergoing a small amount of training, N new exoplanet candidates by
visual inspection of the Kepler lightcurves that were presented in a custom
web-based classification interface. The older Galaxy Zoo morphological
classification project \citep{LintottEtal2008} has also enabled the discovery
of rare objects, via its flexible inspection interface and discussion forum
\citep{LintottEtal2009}. Indeed, several of us (AV,EB,CC,TJ,CM,LW) were active
in an informal Galaxy Zoo gravitational lens search, an experience which led
to the present hypothesis that a systematic online visual lens search could be
successful.

In this paper, we describe the \sw website, an online system that enables 
crowd-sourced gravitational lens detection by inviting volunteers to classify
astronomical survey images as containing lens candidates or not.  In a
companion paper we will present the new gravitational lenses discovered in our
first experimental lens search, and begin to investigate the differences
between lens detections made in \sw and those made with automated techniques.
Here though, we try to answer the following questions:

@@

This paper is organised as follows.  In \Sref{sec:design} we introduce the \sw
classification interface and the volunteers who make up the \sw
collaboration,  explain how we use training images, and describe our two stage
candidate selection strategy. We then briefly introduce, in \Sref{sec:data}
the particular dataset used in the first experimental tests of the \sw system,
and how we prepared the images prior to displaying them in the web interface.
In \Sref{sec:swap} we describe our methodology for interpreting the
classifications made by the volunteers, and then present the results of system
performance tests  made on the training images in \Sref{sec:results}. % In
this section, we also investigate the properties of the crowd, to % understand
where the information is coming from. We discuss the implications of our
results for future lens searches in \Sref{sec:discuss} and draw conclusions in
\Sref{sec:conclude}.

@@

The basic steps of a visual search for gravitational lenses are: 1) prepare
images, 2) display them to an inspector, 3) record the inspector's
classification of each image (as, for example, containing a lens candidate or
not) and 4) analyzing those classifications (and all others) in order to
produce a final candidate list. We describe step 1 in \Sref{sec:data} and step
4 in \Sref{sec:swap}. In this section we take a volunteer's eye view and
begin by describing the \sw classification interface, the crowd of volunteers,
and the interactions between the two.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}
\centerline{
\includegraphics[width=0.9\linewidth]{sw-system-figs/sw-screengrab-marker+feedback.png}}
\caption{Screenshot of the \sw classification interface.}
\label{fig:interface}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Classification Interface}
\label{sec:design:interface}

A screenshot of the \sw classification interface (CI) is shown in
\Fref{fig:interface}. The CI is the centrepiece of the \sw website,
\texttt{http://spacewarps.org}; the web application is written in
coffeescript, css and html and follows the general design of others written by
the Zooniverse team.\footnote{The \sw web application code is open source and
can be accessed from \texttt{https://github.com/zooniverse/Lens-Zoo}}  The
focus of the CI is a large display of the current pre-prepared PNG image of
the ``subject'' being inspected.   When the image is clicked on by the
volunteer, a marker symbol appears where the pointer was. Several markers can
be placed.  The next image moves rapidly in from a queue formed at the right
hand side of the screen when the ``Finished marking'' button is pressed. At
the same time, the positions of the markers are written out to the
classification database, in an entry that also stores the ID of the subject,
the username (or IP address) of the volunteer, a timestamp and some other
metadata.

Gravitational lenses are rare: typically, most of the images will not contain
a lens candidate, and these need to be quickly rejected by the inspector. The
queue allows several images to be pre-loaded while the volunteer is
classifying the current subject, and the rapid movement is designed to
encourage volunteers to classify rapidly.

For the more interesting subjects, the CI offers two features that  enable
further investigation of the subjects. First is the ``Quick Dashboard,'' a
more advanced image viewer. This allows the viewer to compare three different
contrast and color balance settings, to help bring out subtle features, and to
zoom in on interesting regions of the image to assess small features. Markers
can be placed in the Quick Dashboard just the same as in the main CI image
viewer. The second is a link to that subject's page in the project discussion
forum, \texttt{http://talk.spacewarps.org}. Here, volunteers can discuss the
features they have seen either before they submit their classification, or
after, if they ``favorite'' the subject. There is no ``back'' button, however:
each volunteer may only classify a given subject once.

@@

Gravitational lenses are unfamiliar objects to volunteers who are new
to the site. New volunteers need to learn what lenses look like as quickly as
possible, so that they can contribute informative classifications. They also
need to learn what lenses do not look like, in order to reduce the false
positive detection rate.
